Measurement Bias from Address Aliasing

### 2014-05-06 paper.tex
Outline for heap alias section

\section{Bias from Heap Allocation}
Address aliasing can be caused by conflicting pairs of load/store operations to any part of memory.
The previous section looked specifically at an example of collision between stack variables and static data.
Another scenario to consider is collisions in heap allocated memory.
In particular, code that operate on pairs of contiguous arrays can be vulnerable to 4K aliasing.
Acquiring dynamic memory at run time is usually done by calling malloc, which takes a number of bytes to allocate and returns a pointer to that area.
A typical malloc implementation uses two different strategies for how to allocate memory.
The \emph{heap} area in virtual memory is used for smaller allocations.
Programs are initially given a small heap area, which the allocator can increase by calling brk or sbrk.
For larger allocations, it is typically faster to use anonymous memory mapping by calling mmap, instead of growing the ``normal'' heap.
The heap starts at a low address close to the data section and grows towards higher addresses. 
Mmap allocates much closer to the stack.
While heap addresses can look like 0x16e30a0 or 0x1723020, pointers returned by mmap can for example be 0x7f0318a8f010 or 0x7f03105d2010.
This distinction is of course unimportant for application developers, as everything is conceptually the same ``heap''.
However, mmap has an interesting property in that allocations will always be page aligned.
The page size is 4096 bytes, meaning two pointers returned by mmap will \emph{always} alias.\footnote{libc's version of malloc adds 16 bytes of medatada at the beginning, therefore every mmap'ed address ends with 0x010}

@@

It is perhaps unfair to assume the compiler can fix this when only given a generic optimization flag, as aliasing effects is (presumably) an Intel-specific issue.
You might also want to utilize the C99 keyword restrict to explicitly tell the compiler that the buffers are disjoint in memory.

>>>>

\section{Bias from Heap Allocation}
% Intro
% - Bias from aliasing caused by accesses to any part of memory
% - Dynamic memory usually from dynamically linked libs = bias
% - More interesting: default behavior provokes aliasing 
% Heap allocation alias by default
% - Normal heap with brk/sbrk syscalls
%   - Manage heap and increase on demand
% - Different depending on size (and other parameters)
%   - Use mmap for performance, usually large allocations
%   - Posix function, anonymous memory mapping
%   - Page aligned
% - Normal addr. vs mmap addr.
% - Makes no difference ++ everything is conceptually the same heap, but alias by default!
% Aligned sequential access
% - Intel warning, copy routine. 
% - Trouble with aligned read/write buffers, creating constant conflicts in sliding window.
% - More contrived example; convolution
%   - Simple code disregarding endpoints
%   - Manipulate start addresses by offseting pointers
%   - Compiled with gcc O3, shows bad performance (graph) + discussion
% - Table with lots of data points, showing actual numbers for each offset
% Architectural optimization
% - Modified example with restrict and -march=i7
% - Other compilers: Clang (pretty easy), Icc (already used preview version I think...)
% Bahavior of other Allocators
% - Intro each allocator
% - Present table
% - Conclude; not good...

Address aliasing can be caused by conflicting pairs of load and store operations to any part of memory.
The previous section looked specifically at an example of collision between stack variables and static data.
Another scenario to consider is collisions in heap allocated memory.
In particular, code that operate on pairs of contiguous arrays can be vulnerable to 4K aliasing.
Acquiring dynamic memory at run time is usually done by calling malloc, which takes a number of bytes to allocate and returns a pointer to that area.

\subsection{\texttt{malloc} Alias by Default}
A typical malloc implementation uses two different strategies for how to allocate memory.
The \emph{heap} area in virtual memory is used for smaller allocations.
Programs are initially given a small heap area, which the allocator can increase by calling brk or sbrk.
For larger allocations, it is typically faster to use anonymous memory mapping by calling mmap, instead of growing the ``normal'' heap.
The heap starts at a low address close to the data section and grows towards higher addresses. 
Mmap allocates much closer to the stack.
While heap addresses can look like 0x16e30a0 or 0x1723020, pointers returned by mmap can for example be 0x7f0318a8f010 or 0x7f03105d2010.
This distinction is of course unimportant for application developers, as everything is conceptually the same ``heap''.
However, mmap has an interesting property in that allocations will always be page aligned.
The page size is 4096 bytes, meaning two pointers returned by mmap will \emph{always} alias.\footnote{libc's version of malloc adds 16 bytes of medatada at the beginning, therefore every mmap'ed address ends with 0x010}

@@

\subsection{Architectural Optimization}
It is perhaps unfair to assume the compiler can fix this when only given a generic optimization flag, as aliasing effects is (presumably) an Intel-specific issue.
You might also want to utilize the C99 keyword restrict to explicitly tell the compiler that the buffers are disjoint in memory.
